Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-02-05 ICML-submission/appendix-proofs-of-polynomial-approximation-theorems.tex
a pass over the paper

We prove third property by induction on $n$. Indeed, by definition
$$
T_0(\cos(\theta)) = 1 = \cos(0 \theta) \quad \text{and} \quad T_1(\cos(\theta)) = \cos(\theta) \; .
$$
For $n \ge 1$, we have
\begin{align*}
T_{n+1}(\cos(\theta))
& = 2 \cos(\theta) T_n(\cos(\theta)) - T_{n-1}(\cos(\theta)) \\
& = 2 \cos(\theta) \cos(n \theta) - \cos((n-1)\theta)) \; ,
\end{align*}
where the last step follow by induction hypothesis.
It remains to show that the last expression equals $\cos((n+1)\theta)$.
This can be derived from the trigonometric formula
$$
\cos(\alpha \pm \beta) = \cos(\alpha) \cos(\beta) \mp \sin(\alpha) \sin(\beta) \; .
$$
By substituting $\alpha = n \theta$ and $\beta = \theta$, we get two equations
\begin{align*}
\cos((n+1) \theta) & = \cos(n \theta) \cos(\theta) - \sin(n \theta) \sin(\theta) \; , \\
\cos((n-1) \theta) & = \cos(n \theta) \cos(\theta) + \sin(n \theta) \sin(\theta) \; .
\end{align*}
Summing them yields
$$
\cos((n+1)\theta) + \cos((n-1) \theta) = 2 \cos(n \theta) \cos(\theta)
$$
which finishes the proof.

Fourth property has the similar proof as third property. It suffices
to replace $\cos$ and $\sin$ with $\cosh$ and $\sinh$ respectively.

Fifth property follows from the third property. Indeed, for any $z \in [-1,1]$
there exists $\theta \in \R$ such that $\cos \theta = z$. Thus, $|T_n(z)| =
|T_n(\cos(\theta))| = |\cos(n\theta)| \le 1$.

@@

We verify the seventh property by induction on $n$.
For $n=0$ and $n=1$ the inequality trivially holds, since $\norm{T_0} = \norm{T_1} = 1$.
For $n \ge 1$, since $T_{n+1}(z) = 2zT_n(z) - T_{n-1}(z)$,
\begin{align*}
\norm{T_{n+1}}
& \le 2 \norm{T_n} + \norm{T_{n-1}} \\
& \le 2 (1 + \sqrt{2})^n + (1 + \sqrt{2})^{n-1} \\
& = (1 + \sqrt{2})^{n-1} (2 (1 + \sqrt{2}) + 1) \\
& = (1 + \sqrt{2})^{n-1} (3 + 2\sqrt{2}) \\
& = (1 + \sqrt{2})^{n-1} (1 + \sqrt{2})^2 \\
& = (1 + \sqrt{2})^{n+1} \; .
\end{align*}

\end{proof}

Let $r = \left\lceil \log_2(2m) \right\rceil$ and $s = \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil$.
We define the polynomial $p:\R^d \to \R$ as
$$
p(x) = m + \frac{1}{2} - \sum_{i=1}^m \left( T_s(1 - \ip{v_i}{x}) \right)^r \; .
$$
It remains to show that $p$ has properties 1--5.

@@

The third property follows from observation that degree of $p$
is the same as the degree of any one of the terms
$\left( T_s(1 - \ip{v_i}{x}) \right)^r$ which is $r \cdot s$.

To prove the fourth property, we need to upper bound the norm of $p$.
Let $f_i(x) = 1 - \ip{v_i}{x}$, let $g_i(x) = T_s(1 - \ip{v_i}{x})$
and let $h_i(x) = (T_s(1 - \ip{v_i}{x}))^r$. We have
$$
\norm{f_i}^2 = 1 + \norm{v_i}^2 \le 1 + 1 = 2 \; .
$$
Let $T_s(z) = \sum_{j=0}^s c_j z^j$ be the expansion of $s$-th Chebyshev polynomial.
Then,
\begingroup
\allowdisplaybreaks
\begin{align*}
\norm{g_i}^2
& = \norm{ \sum_{j=0}^s c_j (f_i)^j }^2 \\
& \le (s + 1) \sum_{j=0}^s \norm{c_j (f_i)^j}^2 \quad \text{(by~part 3 of \autoref{lemma:properties-of-norm-of-polynomials})} \\
& = (s + 1) \sum_{j=0}^s (c_j)^2 \norm{(f_i)^j}^2 \\
& \le (s + 1) \sum_{j=0}^s (c_j)^2 j^j \norm{f_i}^{2j} \quad \text{(by~part 2 of \autoref{lemma:properties-of-norm-of-polynomials})} \\
& \le (s + 1) \sum_{j=0}^s (c_j)^2 j^j 2^{2j} \\
& \le (s + 1) s^s 2^{2s} \sum_{j=0}^s (c_j)^2 \\
& = (s + 1) s^s 2^{2s} \norm{T_s}^2 \\
& = (s + 1) s^s 2^{2s} (1 + \sqrt{2})^{2s} \quad \text{(by~part 7 of \autoref{proposition:properties-of-chebyshev-polynomials})} \\
& = (s + 1) \left(4(1+\sqrt{2})^2 s \right)^s \\
& \le \left(8(1+\sqrt{2})^2 s \right)^s
& \le \left(47 s \right)^s \; .
\end{align*}
\endgroup
where we used that $s+1 \le 2^s$ for any non-negative integer $s$.
Finally,
\begin{align*}
\norm{p}
& \le m + \frac{1}{2} + \sum_{i=1}^m \norm{(g_i)^r} \\
& = m + \frac{1}{2} + \sum_{i=1}^m \sqrt{\norm{(g_i)^r}^2} \\
& \le m + \frac{1}{2} + \sum_{i=1}^m \sqrt{r^{rs} \norm{g_i}^{2r}} \\
& \le m + \frac{1}{2} + m r^{rs/2} \left(47 s \right)^{rs/2} \\
& = m + \frac{1}{2} + m \left(47 rs \right)^{rs/2} \; .
\end{align*}
We can further upper bound the last expression by using that $m \le \frac{1}{2} 2^r$.
Since $r,s \ge 1$,
\begin{align*}
\norm{p}
& \le m + \frac{1}{2} + m \left(47 rs \right)^{rs/2} \\
& \le \frac{1}{2} 2^r + \frac{1}{2} + \frac{1}{2} 2^r \left(47 rs \right)^{rs/2} \\
& \le 2^r + \frac{1}{2} 2^r \left(47 rs \right)^{rs/2} \\
& = 2^r \left(1 + \frac{1}{2} \left(47 rs \right)^{rs/2} \right) \\
& = 2^r \left(47 rs \right)^{rs/2} \\
& \le 4^{rs/2} \left(47 rs \right)^{rs/2} \\
& \le \left(188 rs \right)^{rs/2} \; .
\end{align*}
Substituting for $r$ and $s$ finishes the proof.

@@

To prove the second part, notice that $P_n(z)$ is non-negative and $P_n(-z)$ is
non-positive for any $z \in [1,2^n]$. Now, fix $z \in [1,2^n]$ and let $j \in
\{1,2,\dots,n\}$ be such that $2^{j-1} \le z \le 2^j$. This implies that
$(z+2^{j})^2 \ge 4 (z - 2^j)^2$. We can write $\frac{P_n(z)}{-P_n(-z)}$ as a
product of $n+1$ non-negative fractions
$$
\frac{P_n(z)}{-P_n(-z)}
= \frac{z-1}{z+1} \cdot \frac{(z-2^j)^2}{(z+2^j)^2} \prod_{\substack{i ~:~ 1 \le i \le n \\ i \neq j}} \frac{(z-2^i)^2}{(z+2^i)^2} \; .
$$
The second part follows from the observation that the second fraction is upper
bounded by $1/4$ and all other fractions are upper bounded by $1$.

@@

\begin{lemma}[Properties of $S_{n,r}$ and $B_{n,r}$]
\label{lemma:properties-of-s-n-r}
Let $n,m$ be non-negative integers.
Let $r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1$. Then,
\begin{enumerate}
\item If $z \in [1,2^n]$ then $S_{n,r}(z) \in [1,1+\frac{1}{2m}]$.
\item If $z \in [-2^n, -1]$ then $S_{n,r} \in [-1-\frac{1}{2m}, -1]$.
\item If $z \in [-1,1]$ then $|S_{n,r}(z)| \le 1$.
\item If $z \in [-2^n,2^n]$ then \\ $B_{n,r}(z) \ge \left(1 - \frac{1}{4m+1} \right) 2^{n(n+1)r}$.
\end{enumerate}
\end{lemma}

@@

\begin{enumerate}
\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{lemma:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. Since
$r \ge \frac{1}{2} \log_2(4m+1)$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$S_{n,r}(z) = \frac{1+c}{1-c} \in [1,1 + \frac{1}{2m}]$.

@@

\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. If $z \in [0,1]$, by part
1 of~\autoref{lemma:properties-of-p-n} and the fact that $r$ is odd, $c
\in [-1,0]$, and thus, $S_{n,r}(z) = \frac{1+c}{1-c} \in [0,1]$. Since
$S_{n,r}(z)$ is odd, for $z \in [-1,0]$, $S_{n,r}(z) \in [-1,0]$.

@@

To prove part 2, consider any $x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~
\norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$. Observe that
$\frac{\ip{v_i}{x}}{\gamma} \in [-\frac{1}{\gamma}, \frac{1}{\gamma}]$. Consider
$S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right)$ for any $i \in
\{1,2,\dots,m\}$. Parts 1,2, and 3 of \autoref{lemma:properties-of-s-n-r}
and the fact $1/\gamma \le 2^s$ imply that
$S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right) \le 1 +
\frac{1}{2m}$ for all $i \in \{1,2,\dots,m\}$. By the choice of $x$, there
exists $j \in \{1,2,\dots,m\}$ such that $\ip{v_j}{x} \le - \gamma$. Part 2 of
\autoref{lemma:properties-of-s-n-r} implies that
$S_{s,r}\left(\frac{\ip{v_j}{x}}{\gamma}\right) \in [-1-\frac{1}{2m},-1]$. Thus,
\begin{align*}
Q(x)
& = \left( \sum_{i=1}^m S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right) - \left(m - \frac{1}{2} \right) \\
& = S_{s,r}\left( \frac{\ip{v_j}{x}}{\gamma} \right) + \left( \sum_{\substack{i ~:~ 1 \le i \le m \\ i \neq j}} S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right)  \\
& \qquad - \left(m - \frac{1}{2} \right) \\
& \le -1 + (m-1) \left( 1 + \frac{1}{2m} \right) - \left(m - \frac{1}{2} \right) \\
& \le 1/2 \; .
\end{align*}
\end{proof}

To prove parts 1 and 2 of \autoref{theorem:polynomial-approximation-2} first
note that part 4 of \autoref{lemma:properties-of-s-n-r} implies that for any $x$
such that $\norm{x} \le 1$, $B_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right)$
is positive. Thus $p(x)$ and $Q(x)$ have the same sign on the unit ball.
Consider any $x$ in either
$\displaystyle \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$
or in
$\displaystyle \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$.
\autoref{lemma:properties-of-q} states that $|Q(x)| \ge 1/2$ and the sign
depends on which of the two sets $x$ lies in. Since signs of $Q(x)$ and $p(x)$
are the same, it remains to show that $|p(x)| \ge m^{s^2m/2}$. Indeed,
\begin{align*}
|p(x)|
& = |Q(x)| \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \\
& \ge |Q(x)| \left( 2^{s(s+1)r} \left( 1 - \frac{1}{4m+1} \right) \right)^m \\
& \ge \frac{1}{2} |Q(x)| \cdot 2^{s(s+1)rm} \\
& \ge \frac{1}{4} \cdot 2^{s(s+1)rm} \quad \text{(\autoref{lemma:properties-of-q})} \; .
\end{align*}
where we used that $\left(1-\frac{1}{4m+1}\right)^m \ge e^{-4} \ge 1/2$.

To prove part 3 of \autoref{theorem:polynomial-approximation-2} note that
$\deg(P_s) = 2s+1$. Thus, $\deg(A_{s,r})$ and $\deg(B_{s,r})$ are at most
$(2s+1)r$. Therefore, $\deg(p) \le (2s+1) rm$.

It remains to prove part 4 of \autoref{theorem:polynomial-approximation-2}.
For any $i \in \{0,1,2,\dots,s\}$ and any $v \in \R^d$ such that $\norm{v} \le 1$
define multivariate polynomials
\begin{align*}
f_{i,v}(x) & = \frac{\ip{v}{x}}{\gamma} - 2^i \; , \\
q_v(x) & = P_s \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
a_v(x) & = A_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
b_v(x) & = B_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; .
\end{align*}
Note that
\begin{multline*}
p(x) = \left[ \sum_{i=1}^m a_{v_i}(x) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}(x) \right]  \\
- \left(m - \frac{1}{2} \right) \prod_{j=1}^n b_{v_j}(x) \; .
\end{multline*}
We bound the norms of these polynomials. We have
$$
\norm{f_{i,v}}^2 = \norm{v}^2/\gamma^2 + 2^{2i} \le 2 \cdot 2^s \; .
$$
where we used that $1/\gamma \le 2^s$ and $\norm{v} \le 1$.
Since $q_v(x) = f_{i,v}(\frac{\ip{v}{x}}{\gamma}) \prod_{i=1}^s \left(f_{i,v}(\frac{\ip{v}{x}}{\gamma})\right)^2$,
using part 1 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $q_v$
as
\begin{align*}
\norm{q_v}^2
& \le (2s+1)^{2s+1} \norm{f_{0,v}}^2 \prod_{i=1}^s \norm{f_{i,v}}^4 \\
& \le  (2s+1)^{2s+1} (2 \cdot 2^s)^{2s + 1} \; .
\end{align*}
Using parts 3 and 2 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $a_v$ as
\begin{align*}
\norm{a_v}^2
& \le 2\norm{(q_v)^r}^2 + 2\norm{(q_{-v})^r}^2 \\
& \le 2 r^{r(2s+1)} \norm{q_v}^{2r} + 2 r^{r(2s+1)} \norm{q_{-v}}^{2r} \\
& \le 4 r^{r(2s+1)} \left((2s+1)^{2s+1} (2 \cdot 2^s)^{2s + 1} \right)^{2r} \\
& = 4 \left(2^{2s} r (4s+2)^2 \right)^{(2s+1)r} \; .
\end{align*}
The same upper bound holds for $\norm{b_v}^2$. Finally,
\begin{align*}
& \norm{p} \\
& \le \left[ \sum_{i=1}^m \norm{a_{v_i} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}} \right] + \left(m - \frac{1}{2} \right) \norm{\prod_{j=1}^m b_{v_j}} \\
& \le \left[ \sum_{i=1}^m m^{(s+1/2)rm} \norm{a_{v_i}} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} \norm{b_{v_j}} \right] \\
& \qquad + \left(m - \frac{1}{2} \right) m^{(s+1/2)rm} \prod_{j=1}^m \norm{b_{v_j}} \\
& \le (2m-1/2) m^{(s+1/2)rm} \left(4 \left(2^{2s} r (4s+2)^2 \right)^{(2s+1)r} \right)^{m/2} \\
& = (2m-1/2) 2^m \cdot \left(2^{2s} rm (4s+2)^2 \right)^{(s+1/2)rm} \; .
\end{align*}

>>>>

We prove the third property by induction on $n$. Indeed, by definition
$$
T_0(\cos(\theta)) = 1 = \cos(0 \theta) \quad \text{and} \quad T_1(\cos(\theta)) = \cos(\theta) \; .
$$
For $n \ge 1$, we have
\begin{align*}
T_{n+1}(\cos(\theta))
& = 2 \cos(\theta) T_n(\cos(\theta)) - T_{n-1}(\cos(\theta)) \\
& = 2 \cos(\theta) \cos(n \theta) - \cos((n-1)\theta)) \; ,
\end{align*}
where the last step follow by induction hypothesis.
It remains to show that the last expression equals $\cos((n+1)\theta)$.
This can be derived from the trigonometric formula
$$
\cos(\alpha \pm \beta) = \cos(\alpha) \cos(\beta) \mp \sin(\alpha) \sin(\beta) \; .
$$
By substituting $\alpha = n \theta$ and $\beta = \theta$, we get two equations
\begin{align*}
\cos((n+1) \theta) & = \cos(n \theta) \cos(\theta) - \sin(n \theta) \sin(\theta) \; , \\
\cos((n-1) \theta) & = \cos(n \theta) \cos(\theta) + \sin(n \theta) \sin(\theta) \; .
\end{align*}
Summing them yields
$$
\cos((n+1)\theta) + \cos((n-1) \theta) = 2 \cos(n \theta) \cos(\theta)
$$
which finishes the proof.

The fourth property has the similar proof as the third property. It suffices
to replace $\cos$ and $\sin$ with $\cosh$ and $\sinh$ respectively.

The fifth property follows from the third property. Indeed, for any $z \in [-1,1]$
there exists $\theta \in \R$ such that $\cos \theta = z$. Thus, $|T_n(z)| =
|T_n(\cos(\theta))| = |\cos(n\theta)| \le 1$.

@@

We verify the seventh property by induction on $n$.
For $n=0$ and $n=1$ the inequality trivially holds, since $\norm{T_0} = \norm{T_1} = 1$.
For $n \ge 1$, since $T_{n+1}(z) = 2zT_n(z) - T_{n-1}(z)$,
\begin{align*}
\norm{T_{n+1}}
& \le 2 \norm{T_n} + \norm{T_{n-1}} \\
& \le 2 (1 + \sqrt{2})^n + (1 + \sqrt{2})^{n-1} \\
& = (1 + \sqrt{2})^{n-1} (2 (1 + \sqrt{2}) + 1) \\
& = (1 + \sqrt{2})^{n-1} (3 + 2\sqrt{2}) \\
& = (1 + \sqrt{2})^{n-1} (1 + \sqrt{2})^2 \\
& = (1 + \sqrt{2})^{n+1} \; .
\qedhere
\end{align*}
\end{proof}

We are now ready to prove~\autoref{theorem:polynomial-approximation-1}.
Let $r = \left\lceil \log_2(2m) \right\rceil$ and $s = \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil$.
We define the polynomial $p:\R^d \to \R$ as
$$
p(x) = m + \frac{1}{2} - \sum_{i=1}^m \left( T_s(1 - \ip{v_i}{x}) \right)^r \; .
$$
It remains to show that $p$ has properties 1--5.

@@

The third property follows from the observation that the degree of $p$
is the same as the degree of any one of the terms
$\left( T_s(1 - \ip{v_i}{x}) \right)^r$ which is $r \cdot s$.

To prove the fourth property, we need to upper bound the norm of $p$.
Let $f_i(x) = 1 - \ip{v_i}{x}$, let $g_i(x) = T_s(1 - \ip{v_i}{x})$
and let $h_i(x) = (T_s(1 - \ip{v_i}{x}))^r$. We have
$$
\norm{f_i}^2 = 1 + \norm{v_i}^2 \le 1 + 1 = 2 \; .
$$
Let $T_s(z) = \sum_{j=0}^s c_j z^j$ be the expansion of $s$-th Chebyshev polynomial.
Then,
\begingroup
\allowdisplaybreaks
\begin{align*}
\norm{g_i}^2
& = \norm{ \sum_{j=0}^s c_j (f_i)^j }^2 \\
& \le (s + 1) \sum_{j=0}^s \norm{c_j (f_i)^j}^2 \quad \text{(by~part 3 of \autoref{lemma:properties-of-norm-of-polynomials})} \\
& = (s + 1) \sum_{j=0}^s (c_j)^2 \norm{(f_i)^j}^2 \\
& \le (s + 1) \sum_{j=0}^s (c_j)^2 j^j \norm{f_i}^{2j} \quad \text{(by~part 2 of \autoref{lemma:properties-of-norm-of-polynomials})} \\
& \le (s + 1) \sum_{j=0}^s (c_j)^2 j^j 2^{2j} \\
& \le (s + 1) s^s 2^{2s} \sum_{j=0}^s (c_j)^2 \\
& = (s + 1) s^s 2^{2s} \norm{T_s}^2 \\
& = (s + 1) s^s 2^{2s} (1 + \sqrt{2})^{2s} \quad \text{(by~part 7 of \autoref{proposition:properties-of-chebyshev-polynomials})} \\
& = (s + 1) \left(4(1+\sqrt{2})^2 s \right)^s \\
& \le \left(8(1+\sqrt{2})^2 s \right)^s \\
& \le \left(47 s \right)^s \; .
\end{align*}
\endgroup
where we used that $s+1 \le 2^s$ for any non-negative integer $s$.
Finally,
\begin{align*}
\norm{p}
& \le m + \frac{1}{2} + \sum_{i=1}^m \norm{(g_i)^r} \\
& = m + \frac{1}{2} + \sum_{i=1}^m \sqrt{\norm{(g_i)^r}^2} \\
& \le m + \frac{1}{2} + \sum_{i=1}^m \sqrt{r^{rs} \norm{g_i}^{2r}} \\
& \le m + \frac{1}{2} + m r^{rs/2} \left(47 s \right)^{rs/2} \\
& = m + \frac{1}{2} + m \left(47 rs \right)^{rs/2} \; .
\end{align*}
We can further upper bound the last expression by using that $m \le \frac{1}{2} 2^r$.
Since $r,s \ge 1$,
\begin{align*}
\norm{p}
& \le m + \frac{1}{2} + m \left(47 rs \right)^{rs/2} \\
& \le \frac{1}{2} 2^r + \frac{1}{2} + \frac{1}{2} 2^r \left(47 rs \right)^{rs/2} \\
& \le 2^r + \frac{1}{2} 2^r \left(47 rs \right)^{rs/2} \\
& = 2^r \left(1 + \frac{1}{2} \left(47 rs \right)^{rs/2} \right) \\
& = 2^r \left(47 rs \right)^{rs/2} \\
& \le 4^{rs/2} \left(47 rs \right)^{rs/2} \\
& \le \left(188 rs \right)^{rs/2} \; .
\end{align*}
Substituting for $r$ and $s$ finishes the proof.

@@

To prove the second part, notice that $P_n(z)$ is non-negative and $P_n(-z)$ is
non-positive for any $z \in [1,2^n]$. Now, fix $z \in [1,2^n]$ and let $j \in
\{1,2,\dots,n\}$ be such that $2^{j-1} \le z \le 2^j$. This implies that
$(z+2^{j})^2 \ge (2^j)^2 \ge 4 (z - 2^j)^2$. We can write $\frac{P_n(z)}{-P_n(-z)}$ as a
product of $n+1$ non-negative fractions
$$
\frac{P_n(z)}{-P_n(-z)}
= \frac{z-1}{z+1} \cdot \frac{(z-2^j)^2}{(z+2^j)^2} \prod_{\substack{i ~:~ 1 \le i \le n \\ i \neq j}} \frac{(z-2^i)^2}{(z+2^i)^2} \; .
$$
The second part follows from the observation that the second fraction is upper
bounded by $1/4$ and all other fractions are upper bounded by $1$.

@@

\begin{lemma}[Properties of $S_{n,r}$ and $B_{n,r}$]
\label{lemma:properties-of-s-n-r}
Let $n,m$ be non-negative integers.
Let $r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1$. Then,
\begin{enumerate}
\item If $z \in [1,2^n]$ then $S_{n,r}(z) \in [1,1+\frac{1}{2m}]$.
\item If $z \in [-2^n, -1]$ then $S_{n,r}(z) \in [-1-\frac{1}{2m}, -1]$.
\item If $z \in [-1,1]$ then $|S_{n,r}(z)| \le 1$.
\item If $z \in [-2^n,2^n]$ then \\ $B_{n,r}(z) \ge \left(1 - \frac{1}{4m+1} \right) 2^{n(n+1)r}$.
\end{enumerate}
\end{lemma}

@@

\begin{enumerate}
\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{lemma:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. Since
$r \ge \frac{1}{2} \log_2(4m+1)$, this means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$S_{n,r}(z) = \frac{1+c}{1-c} \in [1,1 + \frac{1}{2m}]$.

@@

\item Recall that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. If $z \in [0,1]$, by part
1 of~\autoref{lemma:properties-of-p-n} and the fact that $r$ is odd, $c
\in [-1,0]$, and thus, $S_{n,r}(z) = \frac{1+c}{1-c} \in [0,1]$. Since
$S_{n,r}(z)$ is odd, for $z \in [-1,0]$, $S_{n,r}(z) \in [-1,0]$.

%\frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} =

@@

To prove part 2, consider any $x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~
\norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$. Observe that
$\frac{\ip{v_i}{x}}{\gamma} \in [-\frac{1}{\gamma}, \frac{1}{\gamma}]$. Consider
$S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right)$ for any $i \in
\{1,2,\dots,m\}$. Parts 1,2, and 3 of \autoref{lemma:properties-of-s-n-r}
and the fact $1/\gamma \le 2^s$ imply that
$S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right) \le 1 +
\frac{1}{2m}$ for all $i \in \{1,2,\dots,m\}$. By the choice of $x$, there
exists $j \in \{1,2,\dots,m\}$ such that $\ip{v_j}{x} \le - \gamma$. Part 2 of
\autoref{lemma:properties-of-s-n-r} implies that
$S_{s,r}\left(\frac{\ip{v_j}{x}}{\gamma}\right) \in [-1-\frac{1}{2m},-1]$. Thus,
\begin{align*}
Q(x)
& = \left( \sum_{i=1}^m S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right) - \left(m - \frac{1}{2} \right) \\
& = S_{s,r}\left( \frac{\ip{v_j}{x}}{\gamma} \right) + \left( \sum_{\substack{i ~:~ 1 \le i \le m \\ i \neq j}} S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right)  \\
& \qquad - \left(m - \frac{1}{2} \right) \\
& \le -1 + (m-1) \left( 1 + \frac{1}{2m} \right) - \left(m - \frac{1}{2} \right) \\
& \le -1/2 \; .
\qedhere
\end{align*}
\end{proof}

To prove parts 1 and 2 of \autoref{lemma:polynomial-approximation-2} first
note that part 4 of \autoref{lemma:properties-of-s-n-r} implies that for any $x$
such that $\norm{x} \le 1$, $B_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right)$
is positive. Thus $p(x)$ and $Q(x)$ have the same sign on the unit ball.
Consider any $x$ in either
$\displaystyle \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$
or in
$\displaystyle \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$.
\autoref{lemma:properties-of-q} states that $|Q(x)| \ge 1/2$ and the sign
depends on which of the two sets $x$ lies in. Since signs of $Q(x)$ and $p(x)$
are the same, it remains to show that $|p(x)| \ge \frac{1}{4} \cdot 2^{s(s+1)rm}$.
Indeed,
\begin{align*}
|p(x)|
& = |Q(x)| \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \\
& \ge |Q(x)| \left( 2^{s(s+1)r} \left( 1 - \frac{1}{4m+1} \right) \right)^m \\
& \ge \frac{1}{2} |Q(x)| \cdot 2^{s(s+1)rm} \\
& \ge \frac{1}{4} \cdot 2^{s(s+1)rm} \quad \text{(\autoref{lemma:properties-of-q})} \; .
\end{align*}
where we used that $\left(1-\frac{1}{4m+1}\right)^m \ge e^{-\frac 1 4} \ge 1/2$.

%m^{s^2m/2}

To prove part 3 of \autoref{lemma:polynomial-approximation-2} note that
$\deg(P_s) = 2s+1$. Thus, $\deg(A_{s,r})$ and $\deg(B_{s,r})$ are at most
$(2s+1)r$. Therefore, $\deg(p) \le (2s+1) rm$.

It remains to prove part 4 of \autoref{lemma:polynomial-approximation-2}.
For any $i \in \{0,1,2,\dots,s\}$ and any $v \in \R^d$ such that $\norm{v} \le 1$
define multivariate polynomials
\begin{align*}
f_{i,v}(x) & = \frac{\ip{v}{x}}{\gamma} - 2^i \; , \\
q_v(x) & = P_s \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
a_v(x) & = A_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
b_v(x) & = B_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; .
\end{align*}
Note that
\begin{multline*}
p(x) = \left[ \sum_{i=1}^m a_{v_i}(x) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}(x) \right]  \\
- \left(m - \frac{1}{2} \right) \prod_{j=1}^n b_{v_j}(x) \; .
\end{multline*}
We bound the norms of these polynomials. We have
$$
\norm{f_{i,v}}^2 = \norm{v}^2/\gamma^2 + 2^{2i} \le 2 \cdot 2^{2s} \; .
$$
where we used that $1/\gamma \le 2^s$ and $\norm{v} \le 1$.
Since $q_v(x) = f_{i,v}(\frac{\ip{v}{x}}{\gamma}) \prod_{i=1}^s \left(f_{i,v}(\frac{\ip{v}{x}}{\gamma})\right)^2$,
using part 1 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $q_v$
as
\begin{align*}
\norm{q_v}^2
& \le (2s+1)^{2s+1} \norm{f_{0,v}}^2 \prod_{i=1}^s \norm{f_{i,v}}^4 \\
& \le  (2s+1)^{2s+1} (2 \cdot 2^{2s})^{2s + 1} \; .
\end{align*}
Using parts 3 and 2 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $a_v$ as
\begin{align*}
\norm{a_v}^2
& \le 2\norm{(q_v)^r}^2 + 2\norm{(q_{-v})^r}^2 \\
& \le 2 r^{r(2s+1)} (\norm{q_v}^{2})^r + 2 r^{r(2s+1)} (\norm{q_{-v}}^{2})^r \\
& \le 4 r^{r(2s+1)} \left((2s+1)^{2s+1} (2 \cdot 2^s)^{2s + 1} \right)^{r} \\
& = 4 \left(2^{2s} r (4s+2) \right)^{(2s+1)r} \; .
\end{align*}
The same upper bound holds for $\norm{b_v}^2$. Finally,
\begin{align*}
& \norm{p} \\
& \le \left[ \sum_{i=1}^m \norm{a_{v_i} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}} \right] + \left(m - \frac{1}{2} \right) \norm{\prod_{j=1}^m b_{v_j}} \\
& \le \left[ \sum_{i=1}^m m^{(s+1/2)rm} \norm{a_{v_i}} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} \norm{b_{v_j}} \right] \\
& \qquad + \left(m - \frac{1}{2} \right) m^{(s+1/2)rm} \prod_{j=1}^m \norm{b_{v_j}} \\
& \le (2m-1/2) m^{(s+1/2)rm} \left(4 \left(2^{2s} r (4s+2) \right)^{(2s+1)r} \right)^{m/2} \\
& = (2m-1/2) 2^m \cdot \left(2^{2s} rm (4s+2) \right)^{(s+1/2)rm} \; .
\end{align*}
